### 2012-02-12 exam3/1.tex
Touch ups to nat and bool problem.

\item We expect data \emph{and} interface definitions
  (although interfaces may be defined in comments rather than
  with \code{define-interface}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{15}
Last night, Asumu was up late improving the {\tt class} languages.
Unfortunately, he was too tired to test his work, and his last commit
broke the numeric system.

This puts us in the dire position of no longer being able to calculate
the factorial of large numbers, and we turn to you for help.

This is the definition of a class with a {\tt fact} method:

\begin{verbatim}
;; A Fact is (new fact%)
(define-class fact%
  (define (fact n)
    (cond [(zero? n) 1]
          [else (* n (this . fact (sub1 n)))])))
\end{verbatim}

We need to rewrite this method, using the \verb|Natural| interface,
which has the following methods:

\begin{verbatim}
;; A Natural implements:
;; - zero? : -> Boolean
;; is this Natural the zero value?
;; - sub1 : -> Natural
;; produce a Natural one less than this one 
;; (or zero if this one is zero)
;; - times : Natural -> Natural
;; multiply this Natural by the given one
\end{verbatim}

Write a new definition of the \verb|fact%| class using \verb|Natural|,
and \emph{not} using the built-in numbers of the class system or
DrRacket.  You can use the constant \verb|one|, which is a \verb|Natural|
representing $1$.

@@

Now, design a data defintion that implements \verb|Natural|, and
supports the methods \verb|zero?|, \verb|sub1|, and \verb|times|.  Define the
constant \verb|one|.

@@

\begin{verbatim}
;; A Bool implements:
;; - branch : [Action X] [Action X] -> X
;; run the first action if this is true, otherwise run the second action

;; An [Action X] implements:
;; - run : -> X
;; do this action

@@

Following these interface definitions, design data and class
defintions for implementations of the \verb|Bool| interface, and
reimplement the \verb|fact%| class to use them.
You may modify your natural number implementation if neccessary for
your solution.

@@

\begin{solution}
\begin{verbatim}
(define-class fact%
  (define (fact n)
    (n . zero? . branch (new const% 0) (new recur% n))))
(define-class const%
  (fields run))
(define-class recur%
  (fields n)
  (define (run) ((new fact%) . fact (this . n . sub1) . times n)))

>>>>

\item We expect data \emph{and} interface definitions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{15}
Last night, Asumu was up late improving the {\tt class} languages.
Unfortunately, he was too tired to test his work, and his last commit
broke the numeric system.
%
This puts us in the dire position of no longer being able to calculate
the factorial of large numbers, and we turn to you for help.

This is the definition of a class of functional objects
with a {\tt fact} method:

\begin{verbatim}
;; A Fact is (new fact%) and implements 
;; - fact : Natural -> Natural
;;   Compute n!.
(define-class fact%
  (define (fact n)
    (cond [(zero? n) 1]
          [else (* n (this . fact (sub1 n)))])))
\end{verbatim}

Since we can no longer rely on built in numbers to represent
\verb|Natural|s, we will represent \verb|Natural|s with objects and
then rewrite the \verb|fact| method to operate over this new kind of
data.

Here is the \verb|Natural| interface:

\begin{verbatim}
;; A Natural implements:
;; - zero? : -> Boolean
;;   is this Natural the zero value?
;; - sub1 : -> Natural
;;   produce a natural one less than this natural 
;;   (or zero if this natural is zero)
;; - times : Natural -> Natural
;;   multiply this natural by the given natural
\end{verbatim}

\newpage
Write a new definition of the \verb|fact| method using \verb|Natural|,
and \emph{not} using the built-in numbers of the class system or
DrRacket.  You can assume there is a defined constant \verb|one|,
which is a \verb|Natural| object representing $1$.

@@

Now, design a data defintion that implements \verb|Natural|, and
supports the methods \verb|zero?|, \verb|sub1|, and \verb|times|.
Define the constant \verb|one|.  (\emph{Hint}: remember that the
natural numbers can be seen as a recursive union; a natural number is
either zero, or one more than a natural number.)

@@

\begin{verbatim}
;; A Bool implements:
;; - branch : [Action X] [Action X] -> X
;;   run the first action if this bool is true, 
;;   otherwise run the second action

;; An [Action X] implements:
;; - run : -> X
;; run this action

@@

Following these interface definitions, design data and class
defintions for implementations of the \verb|Bool| interface, and
reimplement the \verb|fact| method to use them.
You may modify your natural number implementation if neccessary for
your solution.

@@

\begin{solution}
\begin{verbatim}
(define-class fact%
  (define (fact n)
    (n . zero? . branch (new const% one) (new recur% n))))
(define-class const%
  (fields run))
(define-class recur%
  (fields n)
  (define (run) ((new fact%) . fact (this . n . sub1) . times n)))
